---
title: "Different distances"
author: "Yu"
date: "2024-08-22"
output: html_document
---

```{r setup, include=FALSE}
library(cifti)
library(dbicc)
library(corrplot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(CovTools)
```

## Load full_data_list
```{r}
load("/Users/yu/Library/Mobile Documents/com~apple~CloudDocs/24_Summer/fMRI/02_explore_data/full_data_list.rda")
```

## Define a function to select target session, time points, and ROIs.
```{r}
filter_data = function(data_list, target_sessions, start_time_point, end_time_point, target_ROI = NULL) {
  filtered_data_list = lapply(data_list, function(item) {
    if (item$session %in% target_sessions) {
      data_subset <- item$data[start_time_point:end_time_point, ]
      colnames(data_subset) = item$Parcel
      
      if (!is.null(target_ROI)) {
        data_subset = data_subset[, target_ROI, drop = FALSE]
      }
      
      df = as.data.frame(data_subset)
      df$subject = item$subject
      df$session = item$session
      df = df %>% select(-subject, -session)
      return(df)
    } else {
      return(NULL)
    }
  })
  
  filtered_data_list = Filter(Negate(is.null), filtered_data_list)
  return(filtered_data_list)
}

# Select target session, time points, and ROIs.
target_sessions = sprintf ("ses-func%02d", 1:10)
start_time_point = 1
end_time_point = 818
target_ROI = NULL  # Set particular ROIs if needed: c("R_V1_ROI", "R_MST_ROI", "R_V6_ROI")

full_data_list <- filter_data(data_list, target_sessions, start_time_point, end_time_point, target_ROI)
```

### delete the subject 08
```{r}
filtered_data_list <- full_data_list[-(71:80)]
```

## Look at correlation matrix

### matrix distance computation
```{r}
# mdist function
mdist <- function(datalist, method=c("f")) {
    dmat<-matrix(0,length(datalist),length(datalist))
    
    for (i in 2:length(datalist)) for (j in 1:(i-1)){
        if (method=="r"){
            v1<-datalist[[i]][lower.tri(datalist[[i]], diag = FALSE)]
            v2<-datalist[[j]][lower.tri(datalist[[j]], diag = FALSE)]
            r<-cor(v1,v2)
            dmat[i,j]=dmat[j,i]=sqrt(1-r)
        } else if (method=='l1'){
            dmat[i,j]=dmat[j,i]=sum(abs(datalist[[i]]-datalist[[j]]))
        } else {
            dmat[i,j]=dmat[j,i]=norm(datalist[[i]]-datalist[[j]],type=method)
        }
    }
    return(dmat)
}

# compute correlation matrices
cor_data_list <- lapply(filtered_data_list,cor)

# compute the matrix of distances among the correlation matrices
distmat_l2 <- mdist(cor_data_list)
distmat_l1 <- mdist(cor_data_list, "l1")
distmat_root <- mdist(cor_data_list, "r")
```

## distance plot
```{r,  fig.height = 5, fig.width = 7}
plot_distance_matrix <- function(distmat, title) {
  distmat_df <- as.data.frame(as.table(as.matrix(distmat)))
  names(distmat_df) <- c("Var1", "Var2", "Distance")
  distmat_df$Var2 <- factor(distmat_df$Var2, levels = rev(levels(distmat_df$Var2)))

  colors <- colorRampPalette(c("white", "blue"))(100)

  p <- ggplot(distmat_df, aes(Var1, Var2, fill = Distance)) +
    geom_tile() +
    scale_fill_gradientn(colors = colors) +
    theme_minimal() +
    labs(title = title, x = NULL, y = NULL) +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
          axis.text.y = element_blank(), axis.ticks.y = element_blank())

  return(p)
}

plot_distance_matrix(distmat_l2, "Distance Matrix method l2")
plot_distance_matrix(distmat_l1, "Distance Matrix method l1")
plot_distance_matrix(distmat_root, "Distance Matrix method root")
```

## Compute the reliability  
```{r dbicc}
dm2icc(distmat_l2,9,10)
dm2icc(distmat_l1,9,10)
dm2icc(distmat_root,9,10)
```

### matrix distance computation (using CovDist)
```{r}
## Calculate covariance matrix.
cor_data_list <- lapply(filtered_data_list,cor)

## transform to 3d array
p <- ncol(cor_data_list[[1]])
N <- length(cor_data_list)

cor_array = array(0, dim = c(p, p, N))

for (i in 1:N) {
  cor_array[,,i] <- cor_data_list[[i]]
}

# check positive definite indices
non_positive_definite_indices <- c() 

for (i in seq_len(dim(cor_array)[3])) {
  eigenvalues <- eigen(cor_array[,,i])$values
 
  if (any(eigenvalues <= 0)) {
    non_positive_definite_indices <- c(non_positive_definite_indices, i)
  }
}

print(non_positive_definite_indices) # 33 74 77 82 are not positive definite
```

## adjust to positive definite
```{r}
# define a funtion
adjust_to_positive_definite <- function(matrix) {

    min_eigenvalue <- min(eigen(matrix)$values)
    
    if (min_eigenvalue <= 0) {
        adjustment_value <- abs(min_eigenvalue) + 0.01  # plus 0.01
        matrix <- matrix + diag(adjustment_value, nrow = nrow(matrix))
    }
    return(matrix)
}

# adjust slice 33, 74, 77, 82
non_positive_def_indices <- c(33, 74, 77, 82)
for (i in non_positive_def_indices) {
    cor_array[,,i] <- adjust_to_positive_definite(cor_array[,,i])
}

# check positive definite
for (i in non_positive_def_indices) {
    print(all(eigen(cor_array[,,i])$values > 0))
}
```

## get pairwise distance and reliability

### Euclidean
```{r}
distance_methods <- c("Euclidean")

results <- data.frame(Method = character(), Reliability = numeric(), stringsAsFactors = FALSE)

for (method in distance_methods) {
  dist_matrix_euclidean <- CovDist(cor_array, method = method)
  write.csv(as.data.frame(dist_matrix_euclidean), "dist_matrix_euclidean.csv", row.names = FALSE)
  
  reliability_score_euclidean <- dm2icc(dist_matrix_euclidean, 9, 10)
  
  results_euclidean <- rbind(results, data.frame(Method = method, Reliability = reliability_score_euclidean))
  }
print(results_euclidean)

plot_distance_matrix(dist_matrix_euclidean, "Euclidean") 
```

### Cholesky
```{r}
distance_methods <- c("Cholesky")

results <- data.frame(Method = character(), Reliability = numeric(), stringsAsFactors = FALSE)

for (method in distance_methods) {
  dist_matrix_cholesky <- CovDist(cor_array, method = method)
  
  reliability_score_cholesky <- dm2icc(dist_matrix_cholesky, 9, 10)
  
  results_cholesky <- rbind(results, data.frame(Method = method, Reliability = reliability_score_cholesky))
}

print(results_cholesky)

plot_distance_matrix(dist_matrix_cholesky, "Cholesky")
```

### LERM
```{r}
distance_methods <- c("LERM")

results <- data.frame(Method = character(), Reliability = numeric(), stringsAsFactors = FALSE)

for (method in distance_methods) {
  dist_matrix_lerm <- CovDist(cor_array, method = method)
  
  reliability_score_lerm <- dm2icc(dist_matrix_lerm, 9, 10)
  
  results_lerm <- rbind(results, data.frame(Method = method, Reliability = reliability_score_lerm))
}

print(results_lerm)

plot_distance_matrix(dist_matrix_lerm, "LERM")
```

Cov
### matrix distance computation (using CovDist)
```{r, eval = FALSE}
## Calculate covariance matrix.
cov_data_list <- lapply(filtered_data_list,cov)

## transform to 3d array
p <- ncol(cov_data_list[[1]])
N <- length(cov_data_list)

cov_array = array(0, dim = c(p, p, N))

for (i in 1:N) {
  cov_array[,,i] <- cov_data_list[[i]]
}

# check positive definite indices
non_positive_definite_indices <- c() 

for (i in seq_len(dim(cov_array)[3])) {
  eigenvalues <- eigen(cov_array[,,i])$values
 
  if (any(eigenvalues <= 0)) {
    non_positive_definite_indices <- c(non_positive_definite_indices, i)
  }
}

print(non_positive_definite_indices) # 33 74 77 82 are not positive definite
```

## adjust to positive definite
```{r, eval = FALSE}
# define a funtion
adjust_to_positive_definite <- function(matrix) {

    min_eigenvalue <- min(eigen(matrix)$values)
    
    if (min_eigenvalue <= 0) {
        adjustment_value <- abs(min_eigenvalue) + 0.01  # plus 0.01
        matrix <- matrix + diag(adjustment_value, nrow = nrow(matrix))
    }
    return(matrix)
}

# adjust slice 33, 74, 77, 82
non_positive_def_indices <- c(33, 74, 77, 82)
for (i in non_positive_def_indices) {
    cov_array[,,i] <- adjust_to_positive_definite(cor_array[,,i])
}

# check positive definite
for (i in non_positive_def_indices) {
    print(all(eigen(cov_array[,,i])$values > 0))
}
```

```{r eval = FALSE}
distance_methods <- c("AIRM")

results <- data.frame(Method = character(), Reliability = numeric(), stringsAsFactors = FALSE)

for (method in distance_methods) {
  dist_matrix <- CovDist(cov_array, method = method)
  
  reliability_score <- dm2icc(dist_matrix, 9, 10)
  
  results <- rbind(results, data.frame(Method = method, Reliability = reliability_score))
}

print(results)
```
"AIRM", "Bhattacharyya", "Cholesky", "Euclidean", "Hellinger", "JBLD", "KLDM", "LERM", "Procrustes.SS", "Procrustes.Full", "PowerEuclidean", "RootEuclidean"

```{r}
# Method         Reliability(Cor)   Reliability(Cov)
# AIRM            0.1825003         0.1974096
# Bhattacharyya   NaN
# Cholesky        0.3621463         0.4015352
# Euclidean       0.4538179         0.5831219
# Hellinger       NaN
# JBLD            NaN
# KLDM            0.6667029
# LERM            0.220661          0.2244307
# Procrustes.SS   0.4242606
# Procrustes.Full 0.4571571
# PowerEuclidean  Error
# RootEuclidean   0.4579008
```

## Lag-dependent distance-based ICC.

```{r, echo = FALSE}
# Input:
# dmat  - an n x n distance matrix
# sub   - a vector of length n of subject ID's 
# time  - a vector of length n of times
# 
# Output:
# A list consisting of
# iccfunc - a function mapping Delta to ICC(Delta), where Delta = time lag
# model   - the mgcv::gam model fitted for W(Delta)
# pvalue  - p-value for lag dependence, from the above model
# range   - range of time lags in the data

licc <- function(dmat, sub, time) {
    require(mgcv)
    d2mat <- as.matrix(dmat)^2
    wmask <- outer(sub,sub,"==")
    bmask <- 1 - wmask
    ltmask <- 1*lower.tri(dmat)
    wmat <- d2mat * wmask * ltmask
    bmat <- d2mat * bmask * ltmask
    deltat <- as.matrix(dist(time)) * ltmask
    wmat[wmat==0] <- bmat[bmat==0] <- deltat[deltat==0] <- NA
    # list(wmat,bmat)
    w2 <- as.vector(wmat)
    delta <- as.vector(deltat)
    mod <<- gam(w2 ~ s(delta, k = 9), method="REML")
    func <- Vectorize(function(del) 1 - predict(mod, newdata=data.frame(delta=del)) / mean(bmat, na.rm=TRUE))
    lst <- list(iccfunc = func, model = mod, 
                pvalue=summary(mod)$s.table[1,4], range=range(deltat, na.rm=TRUE))
    class(lst) <- "licc"
    lst
}

# Plot method
plot.licc <- function(x, type="icc", labs=NULL, add.p=TRUE, ...) {
    subtitle <- if (add.p) paste("Lag dependence p =", signif(x$pvalue,3)) else NULL
    if (type=="w") {
        if (is.null(labs)) {
            xlab <- expression(paste("Time lag ", Delta))
            ylab <- "Expected within-subject distance"
        }
        plot(x$model, shift=x$model$coef[1], xlab=xlab, ylab=ylab, sub=subtitle, ...)
    }
    else if (type=="icc") {
        if (is.null(labs)) {
            xlab <- expression(paste("Time lag ", Delta))
            ylab <- "ICC"
        }
        grid <- seq(db$range[1],db$range[2],,401)
        plot(grid,db$iccfunc(grid), type='l', xlab=xlab, ylab=ylab, sub=subtitle, ...)
    }
   
}
```

```{r}
# MSC data with (conventional) ICC that (on average) decreases with time lag

subjects <- c(1:7, 9, 10)
svec <- rep(subjects, each=10)
tvec <- rep(seq(0, by = 1, length.out = 10), times = length(subjects))

# distance l2
par(mfrow=c(1,3), oma = c(0, 0, 3, 0))
require(fields)
image.plot(distmat_l2[,90:1], col=viridis(256), main="Distance matrix_l2")
db <- licc(distmat_l2, svec, tvec)
plot(db, type="w", main=expression(W(Delta)), add.p=FALSE)
plot(db, type="icc", main=expression(ICC(Delta)))
mtext("distance l2 (square root of sum of squared differences)", outer = TRUE, cex = 1.5, line = 1)

# distance l1
par(mfrow=c(1,3), oma = c(0, 0, 3, 0))
require(fields)
image.plot(distmat_l1[,90:1], col=viridis(256), main="Distance matrix_l1")
db <- licc(distmat_l1, svec, tvec)
plot(db, type="w", main=expression(W(Delta)), add.p=FALSE)
plot(db, type="icc", main=expression(ICC(Delta)))
mtext("distance l1 (sum of absolute differences)", outer = TRUE, cex = 1.5, line = 1)

# distance root
par(mfrow=c(1,3), oma = c(0, 0, 3, 0))
require(fields)
image.plot(distmat_root[,90:1], col=viridis(256), main="Distance matrix_root")
db <- licc(distmat_root, svec, tvec)
plot(db, type="w", main=expression(W(Delta)), add.p=FALSE)
plot(db, type="icc", main=expression(ICC(Delta)))
mtext("distance square root of 1 - r", outer = TRUE, cex = 1.5, line = 1)
```

```{r}
# Euclidean
par(mfrow=c(1,3), oma = c(0, 0, 3, 0))
require(fields)
image.plot(dist_matrix_euclidean[,90:1], col=viridis(256), main="Distance matrix_Euclidean")
db <- licc(dist_matrix_euclidean, svec, tvec)
plot(db, type="w", main=expression(W(Delta)), add.p=FALSE)
plot(db, type="icc", main=expression(ICC(Delta)))
mtext("Euclidean", outer = TRUE, cex = 1.5, line = 1)

# Cholesky
par(mfrow=c(1,3), oma = c(0, 0, 3, 0))
require(fields)
image.plot(dist_matrix_cholesky[,90:1], col=viridis(256), main="Distance matrix_Cholesky")
db <- licc(dist_matrix_cholesky, svec, tvec)
plot(db, type="w", main=expression(W(Delta)), add.p=FALSE)
plot(db, type="icc", main=expression(ICC(Delta)))
mtext("Cholesky", outer = TRUE, cex = 1.5, line = 1)

# LERM
par(mfrow=c(1,3), oma = c(0, 0, 3, 0))
require(fields)
image.plot(dist_matrix_lerm[,90:1], col=viridis(256), main="Distance matrix_LERM")
db <- licc(dist_matrix_lerm, svec, tvec)
plot(db, type="w", main=expression(W(Delta)), add.p=FALSE)
plot(db, type="icc", main=expression(ICC(Delta)))
mtext("LERM", outer = TRUE, cex = 1.5, line = 1)
```

