---
title: "Explore the data3"
author: "Yu"
date: "2024-05-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cifti)
library(dbicc)
library(corrplot)
library(dplyr)
library(tidyr)
library(ggplot2)
```


## load full_data_list

```{r}
load("/Users/yu/Desktop/24 Summer/fMRI/02_explore_data/full_data_list.rda")
```


## Select target session, time points, and ROIs.

```{r}
# Define the function.

filter_data = function(data_list, target_sessions, start_time_point, end_time_point, target_ROI = NULL) {
  filtered_data_list = lapply(data_list, function(item) {
    if (item$session %in% target_sessions) {
      data_subset <- item$data[start_time_point:end_time_point, ]
      colnames(data_subset) = item$Parcel
      
      if (!is.null(target_ROI)) {
        data_subset = data_subset[, target_ROI, drop = FALSE]
      }
      
      df = as.data.frame(data_subset)
      df$subject = item$subject
      df$session = item$session
      df = df %>% select(-subject, -session)
      return(df)
    } else {
      return(NULL)
    }
  })
  
  filtered_data_list = Filter(Negate(is.null), filtered_data_list)
  return(filtered_data_list)
}

# Select target session, time points, and ROIs.

target_sessions = c("ses-func01", "ses-func10")
start_time_point = 1
end_time_point = 818
target_ROI = NULL  # Set particular ROIs if needed: c("R_V1_ROI", "R_MST_ROI", "R_V6_ROI")

filtered_data_list <- filter_data(data_list, target_sessions, start_time_point, end_time_point, target_ROI)

```


## Distance matrix computation

```{r mdist}
mdist <- function(datalist, method=c("f")) {
    dmat<-matrix(0,length(datalist),length(datalist))
    
    for (i in 2:length(datalist)) for (j in 1:(i-1)){
        if (method=="r"){
            v1<-datalist[[i]][lower.tri(datalist[[i]], diag = FALSE)]
            v2<-datalist[[j]][lower.tri(datalist[[j]], diag = FALSE)]
            r<-cor(v1,v2)
            dmat[i,j]=dmat[j,i]=sqrt(1-r)
        } else if (method=='l1'){
            dmat[i,j]=dmat[j,i]=sum(abs(datalist[[i]]-datalist[[j]]))
        } else {
            dmat[i,j]=dmat[j,i]=norm(datalist[[i]]-datalist[[j]],type=method)
        }
    }
    return(dmat)
}
```

```{r,  fig.height = 5, fig.width = 7, fig.align = "center"}

# compute correlation matrices
cor_data_list <- lapply(filtered_data_list,cor)

# compute the matrix of distances among the correlation matrices
distmat <- mdist(cor_data_list)

# visualize
plotdmat(distmat, 10, 2)
```


## Compute the reliability  

We can then input the distance matrix to `dm2icc`, to compute the dbICC:
```{r dbicc}
dm2icc(distmat, 10, 2)
```


## Obtain confidence interval for the dbICC, based on 100 boostrap replicates (a larger number of replicates is recommended in practice):

```{r Bootstrap CI}
dm2icc.bt(distmat, 10, 2, nB = 100)
```


## From the plot, No.15 correlation matrix is quite different than others, so I create a plot for it.
The whole plot is too big, so here just show the first 100 ROIs.

```{r}
cor_matrix_15 <- cor_data_list[[15]] 

cor_matrix_15_100 <- cor_matrix_15[1:100, 1:100]

corrplot(cor_matrix_15_100, 
         method = 'color', 
         order = "original", 
         addrect = 2, 
         tl.col = "black", 
         tl.cex = 0.3,
         tl.srt = 45, 
         is.corr = TRUE)
```

## Find some ROIs have high correlations.
```{r}
cor_matrix_15_long = as.data.frame(as.table(cor_matrix_15)) %>% 
  filter(Var1 != Var2 & abs(Freq) > 0.9) %>% 
  arrange(desc(abs(Freq))) 

rois = cor_matrix_15_long %>% 
  count(Var1, sort = TRUE) %>% 
  head(15) %>% 
  pull(Var1) %>% 
  as.factor() %>% 
  droplevels() %>% 
  print()
```


# Draw time series plots for these ROIs
No.15 is the #71 in the full dataset (session 1, subject 8).
No.16 is the #80 in the full dataset (session 10, subject 8).
I select 1-200 time points.

```{r}
data_15 = data_list[[71]] # This is No.15
data_16 = data_list[[80]] # This is No.16

# Get time series dataframe
create_time_series_df = function(data, session_label, end_time_point) {
  df = as.data.frame(data$data)
  colnames(df) = data$Parcel
  df$Time = 1:nrow(df)
  df$Session = session_label
  df %>% pivot_longer(cols = -c(Time, Session), names_to = "ROI", values_to = "Signal") %>%
    filter(ROI %in% rois & Time >= 1 & Time <= end_time_point)
}
```

## Time series plots for time points 1 to 200.

```{r}
end_time_point = 200

df_15 <- create_time_series_df(data_15, "data_15(sub08, ses01)", end_time_point)
df_16 <- create_time_series_df(data_16, "data_16(sub08, ses10)", end_time_point)

# Combine two df
combined_df <- bind_rows(df_15, df_16)

ggplot(combined_df, aes(x = Time, y = Signal, color = Session)) +
  geom_line() +
  facet_wrap(~ ROI, scales = "free_y") +  
  labs(title = "Time Series Plot for Selected ROIs (Time Points 1-200)",
       x = "Time Points",
       y = "Signal Intensity") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Time series plots for time points 1 to 818.

```{r}
end_time_point = 818

df_15 <- create_time_series_df(data_15, "data_15(sub08, ses01)", end_time_point)
df_16 <- create_time_series_df(data_16, "data_16(sub08, ses10)", end_time_point)

# Combine two df
combined_df <- bind_rows(df_15, df_16)

ggplot(combined_df, aes(x = Time, y = Signal, color = Session)) +
  geom_line() +
  facet_wrap(~ ROI, scales = "free_y") +  
  labs(title = "Time Series Plot for Selected ROIs (Time Points 1-818)",
       x = "Time Points",
       y = "Signal Intensity") +
  theme_minimal() +
  theme(legend.position = "bottom")
```